---
title: "Esame scienza di dati - Formula 1 🏎️"
author: "Enrico Franco, 144237"
date: "2023-07-14"
output:
  ioslides_presentation: 
    css: style.css
#output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment="", message = FALSE, warning = FALSE)
```

```{r}
#---Libraries---#

library(dplyr, warn.conflicts =  FALSE)
library(ggplot2)
# label on the plots
library(ggrepel)
# plot side by side
library(patchwork)
# melt dataframe
library(reshape2)
# interactive plot
library(plotly)
# better print table
library(gt)
library(gtExtras)
```

## Formula 1

**Formula 1** is the highest class of international racing for open-wheel single-seater formula racing cars. The FIA Formula 1 World Championship has been one of the premier forms of racing around the world since 1950. 

The word _formula_ refers to the set of rules to which all participants' cars must conform and the season consists of a series of races, known as Grands Prix (GP), in multiple countries and continents around the world.

## Contents

1. Problem and Objective
2. Data preparation
3. Different type of Regressions
4. Finding and conclusion

## Questions

Is it possible to predict the number of total points each driver will finish the Formula 1 Championship with?

# Dataset

## Ergast API

The Ergast Developer API is an experimental web service which provides a historical record of motor racing data for non-commercial purposes. The API provides data for the Formula One series, from the beginning of the world championships in 1950 to today. It's even possible to [download the database tables in CSV format](https://ergast.com/downloads/f1db_csv.zip) or the [SQL image](https://ergast.com/downloads/f1db.sql.gz).

Each CSV file contains a single database table and the first line of each file contains the column headers. 
The tables are described in the [User Guide](https://ergast.com/docs/f1db_user_guide.txt).

---

```{r}
#---Import the files---#

results <- read.csv("f1db_csv/results.csv")

status <- read.csv("f1db_csv/status.csv")

driver_standings <- read.csv("f1db_csv/driver_standings.csv")

drivers <- read.csv("f1db_csv/drivers.csv")
  
constructor_standings <- read.csv("f1db_csv/constructor_standings.csv")

constructors <- read.csv("f1db_csv/constructors.csv")
  
#qualifying <- read.csv("f1db_csv/qualifying.csv")

races <- read.csv("f1db_csv/races.csv")
```

### Results
```{r}
results <- results %>%  
             select(-c(resultId, number, positionText, time, milliseconds, fastestLap, position, rank))

# add status to results
results <- left_join(results, status, by = "statusId")

# convert mean velocity during the fastestlap in number and place 0 instead of NA
results$fastestLapSpeed <- as.numeric(results$fastestLapSpeed) %>% 
  replace(is.na(.), 0)

head(results) %>% 
  gt() %>% 
  gt_theme_espn()
```

### Drivers
```{r}
driver_standings <- driver_standings %>% 
                      select(-c(positionText, driverStandingsId))

# create the name
drivers$driverName <- paste(drivers$forename, drivers$surname, sep=" ")

drivers <- drivers %>% 
             select(-c(driverRef, number, code, nationality, url, forename, surname))

driver_standings <- left_join(driver_standings, drivers, by = "driverId")

head(driver_standings) %>% 
  gt() %>% 
  gt_theme_espn()
```

### Constructors
```{r}
constructor_standings <- constructor_standings %>% 
                            select(-c(constructorStandingsId, positionText))

constructors <- constructors %>% 
                    select(-c(constructorRef, nationality, url))

constructor_standings <- left_join(constructor_standings, constructors, by = "constructorId")

head(constructor_standings) %>% 
  gt() %>% 
  gt_theme_espn()
```

### Races and circuits
```{r}
races <- races %>% 
            select(raceId, year, round, date, circuitId, name)

head(races) %>% 
  gt() %>% 
  gt_theme_espn()
```

## Merging informations to a single dataframe 
```{r}
# add results informations
final_data <- left_join(results, races, by = "raceId") %>% 
                relocate(year, date, circuitId, name, .after = "raceId") #%>% 
                #mutate(date = as.Date(date))
final_data <- rename(final_data, gpName = name)

# add drivers informations
final_data <- left_join(final_data, drivers, by = "driverId") %>% 
                relocate(driverName, dob, .after = "driverId")

# add constructor informations
final_data <- left_join(final_data, constructors, by = "constructorId") %>% 
                relocate(name, .after = "constructorId")
final_data <- rename(final_data, constructorName = name)

head(final_data) %>% 
  gt() %>% 
  gt_theme_espn()
```

## Calculate new features to add to the dataframe

>- **winRate**: Likelihood of winning a race for every driver
```{r}
# count every attended race for each driver
n_driver_races <- final_data %>% 
                    count(driverId)

# count every win for each driver
wins <- final_data %>% 
          filter(positionOrder == 1) %>% 
          count(driverId)

# calculate the rateo in %
win_rate <- inner_join(n_driver_races, wins, by = "driverId") %>%
              mutate(winRate = (n.y / n.x) * 100) %>% 
              select(driverId, winRate)

# add the win rate to the dataframe
final_data <- left_join(final_data, win_rate, by = "driverId")

# driver who never won have 0 instead of NA
final_data$winRate[is.na(final_data$winRate)] <- 0
```

>- **qualiRate**: Likelihood of qualifying first for every driver
```{r}
# count every qualifying round won by each driver
quali_wins <- final_data %>% 
          filter(grid == 1) %>% 
          count(driverId)

# calculate the rateo in %
quali_win_rate <- inner_join(n_driver_races, quali_wins, by = "driverId") %>%
              mutate(qualiRate = (n.y / n.x) * 100) %>% 
              select(driverId, qualiRate)

# add the qualifying rate to the dataframe
final_data <- left_join(final_data, quali_win_rate, by = "driverId") %>% 
                relocate(winRate, qualiRate, .after = "dob")
# driver who never won qualifying have 0 instead of NA
final_data$qualiRate[is.na(final_data$qualiRate)] <- 0
```

```{r}
#---Plot of the Rates just found---#

head(final_data %>% 
  # seelct just one occurence for each driver
  group_by(driverId) %>% 
  slice(1) %>% 
  # arrange the winRate
  arrange(desc(winRate)) %>% 
  ungroup() %>% 
  # select the information to show
  filter(winRate > 0 | qualiRate > 0 ) %>% 
  select(driverName, winRate, qualiRate), n = 15) %>%  # only 15 of them
  # theme
  gt() %>% 
  gt_theme_espn()
```


## Heatmap
```{r}
# select only the numeric data 
heat_final_data <- final_data %>% 
                     select_if(is.numeric)

# create the molten dataframe with the correlation for the heatmap
molten_cor_matrix <- melt(cor(heat_final_data))

ggplot(molten_cor_matrix, aes(x = Var1, y = Var2, fill = value)) + # fill based on the value of the correlation
  geom_tile() +
  # Aesthetic
  scale_fill_gradient2(low = "#ffff99", mid = "#FFcc99", high = "red", name = "Correlation") + # scale of the colors
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + # angle the text on the x axes
  coord_equal() + # make everything a square
  labs(x = "", y = "") + # remove axes' name on the axes
  ggtitle("Heatmap of the dataframe")
```

# Analysis

## Classify races in first and second half of the championship
```{r}
divided_races <- races %>%
  group_by(year) %>%
  arrange(year) %>% 
  mutate(totalRaces = n()) %>% # add number of races to totalRaces
  mutate(halfRaces = ceiling(totalRaces / 2)) %>% # ceiling to add more races the the first half
  slice(1) %>% # filter just the first row for every year
  select(year, totalRaces, halfRaces)

# add the number of total races and half of that to the final dataframe
final_data <- inner_join(divided_races, final_data, by = "year")

# calculate the points for each driver of the first half of championship
first_half_points <- final_data %>%
  group_by(year, driverId) %>%
  filter(round <= halfRaces) %>% # only the races in the first half of the championship
  group_by(year, driverId) %>%
  summarise(firstHalfPoints = sum(points), .groups = "drop") # add the sum on points for each driver and place it in the firstHalfPoints column
```

```{r}
# sum all points for each driver
all_points <- final_data %>% 
               group_by(year, driverId) %>% 
               summarise(allPoints = sum(points), .groups = "drop")

# create the data used to plot the training data
plot_data <- full_join(first_half_points, all_points, by = c("year", "driverId")) %>% 
               replace(is.na(.), 0) # place 0 in the firstHalfPoints where it's NA caused by drivers run just on the second half of the championship

# create the 2023 dataframe
data_2023 <- subset(plot_data, year == 2023) %>% 
               #rename(new_firstHalfPoints = firstHalfPoints) %>% 
               ungroup() 
#remove the 2023 data for the training
plot_data <- subset(plot_data, year != 2023)

ggplot(plot_data, aes(firstHalfPoints, allPoints, colour = driverId)) +
  geom_point(alpha = 0.8) +
  theme(legend.position="none") +
  # labels
  geom_label(data = subset(plot_data, firstHalfPoints == max(firstHalfPoints)), 
             aes(label = drivers$driverName[drivers$driverId == driverId]), 
             # position of the label
             vjust = 1, nudge_x = 6, nudge_y = -10) +
  geom_label(data = subset(plot_data, allPoints == max(allPoints)), 
             aes(label = drivers$driverName[drivers$driverId == driverId]),
             # position of the label
             hjust = 1, nudge_x = -3) +
  # axes
  xlim(0, 245) + # limit of the x axes
  xlab("Points scored in half the season") +
  ylab("Points scored in all the season") +
  # title
  ggtitle("Results points")
```

## Why points are concentrated near the 0?
Let's find out: 

>- the number of points awarded to the 1st place over the years
>- the number of races on the calendar over the years
>- the current point system for every race

## Number of points awarded to the 1st place over the years
```{r}
# select the number of points given to a win during the years
win_score <- final_data %>%
  group_by(year) %>%
  summarise(pointsForWin = max(points), .groups = "drop") %>%
  ungroup()

ggplot(win_score, aes(year, pointsForWin, fill = as.factor(pointsForWin))) +
  # axes
  xlab("Year") +
  ylab("Points for a single win") +
  # title
  ggtitle("Points for a win in all the years") +
  # legend
  scale_fill_discrete(name = "Points for a single Win") +
  geom_bar(stat = "identity") # to use the number as the height of the bars

# removed data from 2014 as it's an outlier and the last seasons (2023)
final_data <- subset(final_data, year != 2014 & year != 2023) 
plot_data <- subset(plot_data, year != 2014)
```

## Number of races run for each championship
```{r}
# Calculate the number of races run for each championship
calendar <- final_data %>%
  group_by(year) %>%
  summarise(racesForYear = max(round), .groups = "drop") %>%
  ungroup()

ggplot(calendar, aes(year, racesForYear, fill = as.factor(racesForYear))) +
  # axes
  xlab("Year") +
  ylab("Races in the season") +
  # title
  ggtitle("Number of races in all the years") +
  # legend
  scale_fill_discrete(name = "Number of races") +
  geom_bar(stat = "identity") # to use the number as the height of the bars
```

## The current point system introduced by the FIA 

```{r}
fia_points <- final_data %>% 
  # filter the last race
  filter(year == 2022 & round == 10 & points > 0) %>% 
  ungroup() %>% 
  # selection
  select(positionOrder, points) %>% 
  rename(position = positionOrder)

# removed the fastest lap point
fia_points[fia_points == 16] <- 15

fia_points %>% 
  # theme
  gt() %>% 
  gt_theme_espn()
```
  
and 1 point is added to the driver who made the fastest lap during the race.

# Training Dataframe

## Calculate: 
>- new points using the last point system introduced by the FIA
>- average points scored by each driver
>- copy the qualifying and win rates to the training dataframe

```{r}
# function to calculate the points using the new point system
calculate_points <- function(data) {
  result <- data %>%
    group_by(driverId, positionOrder, year) %>%
    summarise(number = n(), .groups = "drop") %>% # count the races for each driver in each year
    # create a new column called points 
    mutate(points = case_when( # switch case that change the old points with the new ones
      positionOrder == 1 ~ 25,
      positionOrder == 2 ~ 18,
      positionOrder == 3 ~ 15,
      positionOrder == 4 ~ 12,
      positionOrder == 5 ~ 10,
      positionOrder == 6 ~ 8,
      positionOrder == 7 ~ 6,
      positionOrder == 8 ~ 4,
      positionOrder == 9 ~ 2,
      positionOrder == 10 ~ 1,
      TRUE ~ 0
    )) %>%
    mutate(points_per_driver = number * points) %>% # calculate the new points for the race
    group_by(driverId, year) %>%
    summarise(new_points = sum(points_per_driver), .groups = "drop") %>% # sum every points for all the races a single driver run
    ungroup()
    
  return(result)
}

# convert all the points with the new system
current_all_points <- calculate_points(final_data) %>% 
                        rename(new_allPoints = new_points)

# convert just the points of the first half of championships the points with the new system
current_first_half_points <- final_data %>%
                               filter(round <= halfRaces) %>%
                               calculate_points() %>% 
                               rename(new_firstHalfPoints = new_points) # rename the column
```

```{r}
# inner_join() to remove the third pilots or ones who just run a single race
training_data <- inner_join(plot_data, current_first_half_points, by = c("driverId", "year")) %>% # remove third pilots
  left_join(current_all_points, by = c("driverId", "year")) %>% 
  arrange(driverId) %>% 
  # add the name of the driver
  left_join(drivers[,c("driverId", "driverName")], by = "driverId") %>% 
  relocate(driverName, .after = "driverId")
```

```{r}
# calculate the average points score by each driver
training_data <- training_data %>% 
  group_by(driverId) %>% 
  mutate(averagePoints = mean(allPoints)) %>% # mean of allPoints for every driver
  ungroup()
```

```{r}
# add number of races to the training dataframe
training_data <- left_join(training_data, n_driver_races, by = "driverId") %>% 
  relocate(n, .after = "driverId") %>% 
  rename(races = n)
```

```{r}
# add driver's win rate to the training data
driver_win_rate <- unique(final_data %>% # get just one for each driver
                     group_by(driverId) %>% 
                     select(driverId, winRate))

# add driver's qualifying rate to the training data
driver_quali_rate <- unique(final_data %>% # get just one for each driver
                     group_by(driverId) %>% 
                     select(driverId, qualiRate))

training_data <- left_join(training_data, driver_quali_rate, by = "driverId") %>% 
                    left_join(., driver_win_rate, by = "driverId") %>% 
                    # move the column to make it easier to see
                    relocate(qualiRate, winRate, .after = "races")
```

## Training dataframe

```{r}
#---Print the training dataframe---#
head(training_data %>% 
       # order the dataframe
       arrange(desc(new_allPoints))) %>% 
        # theme
        gt() %>% 
        gt_theme_espn()
```

## Plot of the new point side by side with the old ones {.flexbox .vcenter}

```{r}
# plot the old points
old <- ggplotly(ggplot(training_data, aes(firstHalfPoints, allPoints, colour = driverId, text = driverName)) + 
                  geom_point(alpha = 0.8) +
                  theme(legend.position="none") + # remove the legend
                  # axes
                  xlab("Points scored in half the season") +
                  ylab("Points scored in all the season"), #+
                  # title
                  #ggtitle("Original points system"),
                tooltip = "text")

new <- ggplotly(ggplot(training_data, aes(new_firstHalfPoints, new_allPoints, colour = driverId, text = driverName)) + 
                  geom_point(alpha = 0.8) +
                  theme(legend.position="none") + # remove the legend
                  # axes
                  xlab("Points scored in half the season") +
                  ylab("Points scored in all the season"), #+
                  # title
                  #ggtitle("New points system"),
                tooltip = "text")

# print them side by side
subplot(old, new) %>% 
  # title of the plot
  layout(#title = 'Original vs New points system', 
         # subtitles as title of the single plot
         annotations = list(
           list( 
             # position of the text
             x = 0.25,  
             y = 1.0,  
             text = "Original points system",  
             # reference for position
             xref = "paper",  
             yref = "paper",  
             # place it on top of the graph
             xanchor = "center",  
             yanchor = "bottom",
             # disable the arrow
             showarrow = FALSE), 
           list( 
             # position of the text
             x = 0.75,  
             y = 1, 
             # text
             text = "New points system", 
             # reference for position
             xref = "paper",  
             yref = "paper",  
             # place it on top of the graph
             xanchor = "center",  
             yanchor = "bottom",
             # disable the arrow
             showarrow = FALSE)
           )
         )
```

## Heatmap of the training data

```{r}
# create the molten dataframe with the correlation for the heatmap
molten_cor_training <- melt(cor(training_data %>% 
                                  select_if(is.numeric)
)) 

ggplot(molten_cor_training, aes(x = Var1, y = Var2, fill = value)) + # fill based on the value of the correlation
  geom_tile() +
  scale_fill_gradient2(low = "#ffff99", mid = "#FFcc99", high = "red", name = "Correlation") + # scale of the colors
  theme_minimal() +
  coord_equal() + # make everything a square
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + # angle the text on the x axes
  labs(x = "", y = "") + # remove axes' name on the axes
  # title
  ggtitle("Heatmap of the training data")
```

## Linear Regression

**Linear** regression is a linear approach for modelling the relationship between a scalar response and one or more explanatory variables where the relationships are modeled using **linear predictor functions** whose unknown model parameters are estimated from the data.

### with just half the points

```{r}
#--- Original points ---#
# Calculate linear regression with just the points of the first half of the championship
linear_regression_old <- training_data %>% 
                           lm(allPoints ~ firstHalfPoints, data = .)

# print the summary and information about the linear regression
#paste0("R-squared: ", summary(linear_regression_old)$r.squared)

# plot of the predicted value on top of the original ones using the linear regression
old1 <- ggplot(training_data, aes(firstHalfPoints, allPoints, colour = driverId)) +
  geom_point(alpha = 0.8) +
  # add the predicted points colored red
  geom_point(aes(x = firstHalfPoints, y = predict(linear_regression_old)), color = "red", alpha = 0.8) +
  theme(legend.position="none") + # remove the legend
  # axes
  xlab("Points scored in half the season") +
  ylab("Points scored in all the season") +
  # title
  ggtitle("Linear regression",subtitle = "Original points system")
```

```{r}
#--- New points ---#
# Calculate linear regression with just the points of the first half of the championship
linear_regression_new <- training_data %>% 
                           lm(new_allPoints ~ new_firstHalfPoints, data = .)

# print the summary and information about the linear regressions
#cat("\tOriginal Points \t\t\t\tNew Points")
#cat("\tR-squared: ", summary(linear_regression_old)$r.squared, "\t\t\tR-squared: ", #summary(linear_regression_new)$r.squared)

lm_r1 <- tibble("Original Points" = summary(linear_regression_old)$r.squared,
       "New Points" = summary(linear_regression_new)$r.squared) 

lm_r1 %>% 
  gt %>% 
   tab_header(
    title = "R-squared"
  ) %>% 
  gt_theme_espn()

# plot of the predicted value on top of the new ones using the linear regression
new1 <- ggplot(training_data, aes(new_firstHalfPoints, new_allPoints, colour = driverId)) +
  geom_point(alpha = 0.8) +
  # add the predicted points colored red
  geom_point(aes(x = new_firstHalfPoints, y = predict(linear_regression_new)), color = "red", alpha = 0.8) +
  theme(legend.position="none") + # remove the legend
  # axes
  xlab("Points scored in half the season") +
  ylab("Points scored in all the season") +
  # title
  ggtitle("Linear regression",subtitle = "New points system")

# print them side by side
old1 + new1
```

### with more features

```{r}
#--- Original points ---#

# Calculate linear regression with more features than just the points of the first half of the championship
linear_regression_old2 <- training_data %>% 
                       lm(allPoints ~ firstHalfPoints + qualiRate + winRate + races + averagePoints, data = .)

# print the summary and information about the linear regression
#paste0("R-squared: ", summary(linear_regression_old2)$r.squared)

# plot of the predicted value on top of the original ones using the linear regression
old2 <- ggplot(training_data, aes(firstHalfPoints, allPoints, colour = driverId)) +
  geom_point(alpha = 0.8) +
  # add the predicted points colored red
  geom_point(aes(x = firstHalfPoints, y = predict(linear_regression_old2), alpha = 0.8), color = "red") +
  theme(legend.position="none") + # remove the legend  
  # axes
  xlab("Points scored in half the season") +
  ylab("Points scored in all the season") +
  # title
  ggtitle("Linear regression",subtitle = "Original points system")
```

```{r}
#--- New points ---#

# Calculate linear regression with more features than just the points of the first half of the championship
linear_regression_new2 <- training_data %>% 
                       lm(new_allPoints ~ new_firstHalfPoints + qualiRate + winRate + races + averagePoints, data = .)

# print the summary and information about the linear regression
#paste0("R-squared: ", summary(linear_regression_new2)$r.squared)

lm_r2 <-tibble("Original Points" = summary(linear_regression_old2)$r.squared,
       "New Points" = summary(linear_regression_new2)$r.squared)

lm_r2 %>% 
  gt %>% 
   tab_header(
    title = "R-squared"
  ) %>% 
  gt_theme_espn()

# plot of the predicted value on top of the original ones using the linear regression
new2 <- ggplot(training_data, aes(new_firstHalfPoints, new_allPoints, colour = driverId)) +
  geom_point(alpha = 0.8) +
  # add the predicted points colored red
  geom_point(aes(x = new_firstHalfPoints, y = predict(linear_regression_new2), alpha = 0.8), color = "red") +
  theme(legend.position="none") + # remove the legend
  # axes
  xlab("Points scored in half the season") +
  ylab("Points scored in all the season") +
  # title
  ggtitle("Linear regression",subtitle = "New points system")

# print them side by side
old2 + new2
```

## Polynomial Regression 

**Polynomial** regression is a form of regression analysis where the relationship between the independent variable and the dependent variable is modeled as an ^nth^ degree polynomial in x. It fits a **nonlinear relationship** between the value of x and the corresponding conditional mean of y.

### with one feature

```{r}
#--- Original points ---#

# Calculate polynomial regression with just the points of the first half of the championship
poly_regression_old <- training_data %>% 
                       lm(allPoints ~ poly(firstHalfPoints, 4), data = .)

# print the summary and information about the polynomial regression
#paste0("R-squared: ", summary(poly_regression_old)$r.squared)

# plot of the predicted value on top of the original ones using the polynomial regression
poly_old1 <- ggplot(training_data, aes(firstHalfPoints, allPoints, colour = driverId)) +
  geom_point(alpha = 0.8) +
  # add the predicted points colored red
  geom_point(aes(x = firstHalfPoints, y = predict(poly_regression_old), alpha = 0.8), color = "red") +
  theme(legend.position="none") + # remove the legend
  # axes
  xlab("Points scored in half the season") +
  ylab("Points scored in all the season") +
  # title
  ggtitle("Polynomial regression",subtitle = "Original points system")
```

```{r}
#--- New points ---#

# Calculate polynomial regression with just the points of the first half of the championship
poly_regression_new <- training_data %>% 
                       lm(new_allPoints ~ poly(new_firstHalfPoints, 4), data = .)

# print the summary and information about the polynomial regression
#paste0("R-squared: ", summary(poly_regression_new)$r.squared)
poly_r1 <- tibble("Original Points" = summary(poly_regression_old)$r.squared,
       "New Points" = summary(poly_regression_new)$r.squared)

poly_r1  %>% 
  gt %>% 
   tab_header(
    title = "R-squared"
  ) %>% 
  gt_theme_espn()

# plot of the predicted value on top of the original ones using the polynomial regression
poly_new1 <- ggplot(training_data, aes(new_firstHalfPoints, new_allPoints, colour = driverId)) +
  geom_point(alpha = 0.8) +
  # add the predicted points colored red
  geom_point(aes(x = new_firstHalfPoints, y = predict(poly_regression_new), alpha = 0.8), color = "red") +
  theme(legend.position="none") + # remove the legend
  # axes
  xlab("Points scored in half the season") +
  ylab("Points scored in all the season") +
  # title
  ggtitle("Polynomial regression",subtitle = "New points system")

# print them side by side
poly_old1 + poly_new1
```

### with more features

```{r}
#--- Original points ---#

# Calculate polynomial regression with more features than just the points of the first half of the championship
poly_regression_old2 <- training_data %>% 
                       lm(allPoints ~ poly(firstHalfPoints + qualiRate + winRate, 3), data = .) # removed races and averagePoints

# print the summary and information about the polynomial regression
#paste0("R-squared: ", summary(poly_regression_old2)$r.squared)

# plot of the predicted value on top of the original ones using the polynomial regression
poly_old2 <- ggplot(training_data, aes(firstHalfPoints, allPoints, colour = driverId)) +
  geom_point(alpha = 0.8) +
  # add the predicted points colored red
  geom_point(aes(x = firstHalfPoints, y = predict(poly_regression_old2), alpha = 0.8), color = "red") +
  ggtitle("old points") +
  theme(legend.position="none") +
  # axes
  xlab("Points scored in half the season") +
  ylab("Points scored in all the season") +
  # title
  ggtitle("Polynomial regression",subtitle = "Original points system")
```

```{r}
#--- New points ---#

# Calculate polynomial regression with more features than just the points of the first half of the championship
poly_regression_new2 <- training_data %>% 
                       lm(new_allPoints ~ poly(new_firstHalfPoints + qualiRate + winRate, 3), data = .) # removed races and averagePoints

# print the summary and information about the polynomial regression
#paste0("R-squared: ", summary(poly_regression_new2)$r.squared)
poly_r2 <-tibble("Original Points" = summary(poly_regression_old2)$r.squared,
       "New Points" = summary(poly_regression_new2)$r.squared)

poly_r2 %>% 
  gt %>% 
   tab_header(
    title = "R-squared"
  ) %>% 
  gt_theme_espn()

# plot of the predicted value on top of the original ones using the polynomial regression
poly_new2 <- ggplot(training_data, aes(new_firstHalfPoints, new_allPoints, colour = driverId)) +
  geom_point(alpha = 0.8) +
  # add the predicted points colored red
  geom_point(aes(x = new_firstHalfPoints, y = predict(poly_regression_new2), alpha = 0.8), color = "red") +
  ggtitle("new points") +
  theme(legend.position="none") +
  # axes
  xlab("Points scored in half the season") +
  ylab("Points scored in all the season") +
  # title
  ggtitle("Polynomial regression",subtitle = "New points system")

# print them side by side
poly_old2 + poly_new2
```

## Bayesian linear regression

With the **Bayesian** regression, we formulate linear regression using **probability distributions** rather than point estimates. The response, is not estimated as a single value, but is assumed to be drawn from a probability distribution.

### with one feauture

```{r}
#--- Original points ---#

# import the library for the Bayesian linear regression and analises
library(rstanarm)
library(bayesplot)

# Calculate bayesan regression with more features than just the points of the first half of the championship regression with more features than just the points of the first half of the championship
capture.output(bayes_old1 <- training_data %>% # capture.output() needed to stop printing all the information about the sampling 
  stan_glm(allPoints ~ firstHalfPoints, data = .), file = nullfile())

# print the summary and information about the Bayesian linear regression
#summary(bayes_old1)

# plot of the predicted value on top of the original ones using the quantile regression
old3 <- ggplot(training_data, aes(firstHalfPoints, allPoints, colour = driverId)) +
  geom_point(alpha = 0.8) +
  # add the predicted points colored red
  geom_point(aes(x = firstHalfPoints, y = predict(bayes_old1), alpha = 0.8), color = "red") +
  theme(legend.position="none") + # remove the legend
  # axes
  xlab("Points scored in half the season") +
  ylab("Points scored in all the season") +
  # title
  ggtitle("Bayern's regression",subtitle = "Original points system")
```

```{r}
#--- New points ---#

# Calculate bayesan regression with more features than just the points of the first half of the championship regression with more features than just the points of the first half of the championship
capture.output(bayes_new1 <- training_data %>% # capture.output() needed to stop printing all the information about the sampling 
  stan_glm(new_allPoints ~ new_firstHalfPoints, data = .), file = nullfile())

# print the summary and information about the Bayesian linear regression
#summary(bayes_new2)
bayesan_var1 <- tibble("Original Points" = mean(bayes_R2(bayes_old1)),
       "New Points" = mean(bayes_R2(bayes_new1)))

bayesan_var1 %>% 
  gt %>% 
   tab_header(
    title = "Variance"
  ) %>% 
  gt_theme_espn()

# plot of the predicted value on top of the original ones using the quantile regression
new3 <- ggplot(training_data, aes(new_firstHalfPoints, new_allPoints, colour = driverId)) +
  geom_point(alpha = 0.8) +
  # add the predicted points colored red
  geom_point(aes(x = new_firstHalfPoints, y = predict(bayes_new1), alpha = 0.8), color = "red") +
  theme(legend.position="none") + # remove the legend
  # axes
  xlab("Points scored in half the season") +
  ylab("Points scored in all the season") +
  # title
  ggtitle("Bayern's regression",subtitle = "New points system")

# print them side by side
old3 + new3
```

### with more features

```{r}
#--- Original points ---#

# Calculate bayesan regression with more features than just the points of the first half of the championship regression with more features than just the points of the first half of the championship
capture.output(bayes_old2 <- training_data %>% # capture.output() needed to stop printing all the information about the sampling 
  stan_glm(allPoints ~ firstHalfPoints + qualiRate + winRate + races + averagePoints, data = .), file = nullfile())

# print the summary and information about the Bayesian linear regression
#summary(bayes_old2)

# plot of the predicted value on top of the original ones using the quantile regression
old3 <- ggplot(training_data, aes(firstHalfPoints, allPoints, colour = driverId)) +
  geom_point(alpha = 0.8) +
  # add the predicted points colored red
  geom_point(aes(x = firstHalfPoints, y = predict(bayes_old2), alpha = 0.8), color = "red") +
  theme(legend.position="none") + # remove the legend
  # axes
  xlab("Points scored in half the season") +
  ylab("Points scored in all the season") +
  # title
  ggtitle("Bayern's regression",subtitle = "Original points system")
```

```{r}
#--- New points ---#

# Calculate bayesan regression with more features than just the points of the first half of the championship regression with more features than just the points of the first half of the championship
capture.output(bayes_new2 <- training_data %>% # capture.output() needed to stop printing all the information about the sampling 
  stan_glm(new_allPoints ~ new_firstHalfPoints + qualiRate + winRate + races + averagePoints, data = .), file = nullfile())

# print the summary and information about the Bayesian linear regression
#summary(bayes_new1)
bayesan_var2 <- tibble("Original Points" = mean(bayes_R2(bayes_old2)),
       "New Points" = mean(bayes_R2(bayes_new2)))

bayesan_var2 %>% 
  gt %>% 
   tab_header(
    title = "Variance"
  ) %>% 
  gt_theme_espn()

# plot of the predicted value on top of the original ones using the quantile regression
new3 <- ggplot(training_data, aes(new_firstHalfPoints, new_allPoints, colour = driverId)) +
  geom_point(alpha = 0.8) +
  # add the predicted points colored red
  geom_point(aes(x = new_firstHalfPoints, y = predict(bayes_new2), alpha = 0.8), color = "red") +
  theme(legend.position="none") + # remove the legend
  # axes
  xlab("Points scored in half the season") +
  ylab("Points scored in all the season") +
  # title
  ggtitle("Bayern's regression",subtitle = "New points system")

# print them side by side
old3 + new3
```

# Findings and Conclusion

## Comparison between all the models

Comparison between the _R-squared_ of all the linear and polynomial regressions and using the _variance_ for the bayesan linear regressions divided by original and new points systems.

```{r}
# create the dataframe
comparison <- add_row(lm_r1, lm_r2) %>% 
  add_row(poly_r1) %>% 
  add_row(poly_r2) %>% 
  add_row(bayesan_var1) %>% 
  add_row(bayesan_var2) %>% 
  # add the name of the regression used
  mutate(type = c("Linear Regression", 
                               "Linear Regression with more features", 
                               "Polynomial Regression", 
                               "Polynomial Regression with more features", 
                               "Bayesan Linear Regression",
                               "Bayesan Linear Regression with more features")) %>% 
  # move the names in front 
  relocate(type, .before = "Original Points")

comparison %>% 
  # remove the title from the name column
  gt(rowname_col = "type") %>% 
  # color the row with the max value
  tab_style(style = cell_fill(color = "lightgreen"), # fill the cell with the selected color
    locations = cells_body( # position of the cell
      rows = which(comparison[-1] == max(comparison[-1]), arr.ind = TRUE)[, "row"] # row number
      ))
```



## Predict values for the 2023 season

Using the best suited model, the Linear regression model with more features in this case, and assuming that we are in the second half of the championship we can predict the points in the end will be these:

```{r}
# create a datafame with the name of the driver and the calculated point at the end of the season
predicted_data <- data_2023 %>%
  group_by(driverId) %>% 
  mutate(driverName = drivers$driverName[drivers$driverId == driverId]) %>% 
  ungroup() %>% 
  # calculate the points
  mutate(predicted_points = floor(predict(linear_regression_old, data_2023))) %>%
  # select what to print
  select(driverName, predicted_points) %>% 
  arrange(desc(predicted_points)) %>% 
  rename("Driver Name" = driverName) %>% 
  rename("Predicted points" = predicted_points)

# print predicted values
predicted_data %>% 
  # theme
  gt() %>% 
  gt_theme_espn()
```
